Multivariate Regression

### 2019-02-28 multivariate_regression.Rmd
test plsr on CCLE+CTRP data w/ GE

```{r libraries}
library(tidyverse)
library(pls)
```

@@

There are two component selection strategies in `seelctNcomp`. (1) Is the one-sigma heuristic that consists of choosing the model with the fewest components that is still less than one standard error away from the overall best model. (2) Uses a permutation approach that tests whether adding a new component is beneficial by taking the global minimum of the crossvalidation curve and reducing the number of components one by one as long as no significant deterioration in performance is found (default is $\alpha = 0.01$).
```{r}
par(mfrow = c(1, 2))
ncomp.onesigma <- selectNcomp(gas1, method = "onesigma", plot = TRUE, main = "One-Sigma")
ncomp.permut <- selectNcomp(gas1, method = "randomization", plot = TRUE, main = "Permutation")
par(mfrow = c(1, 1))
```

@@

##

>>>>

```{r libraries}
library(reshape2)
library(tidyverse)
library(pls)
```

@@

There are two component selection strategies in `selectNcomp`. (1) Is the one-sigma heuristic that consists of choosing the model with the fewest components that is still less than one standard error away from the overall best model. (2) Uses a permutation approach that tests whether adding a new component is beneficial by taking the global minimum of the crossvalidation curve and reducing the number of components one by one as long as no significant deterioration in performance is found (default is $\alpha = 0.01$).
```{r}
par(mfrow = c(1, 2))
ncomp.onesigma <- selectNcomp(gas1, method = "onesigma", plot = TRUE, main = "One-Sigma")
ncomp.permut <- selectNcomp(gas1, method = "randomization", plot = TRUE, main = "Permutation")
par(mfrow = c(1, 1))
```

@@

# Test `pls` with drugscreen data

## Global data files

Load in G2P data for gene filtering:
```{r}
ccl_converter <- read.delim("./data_munging/cell_line_database_v1_20180911.tsv", row.names = 1, sep = "\t", header = TRUE)
colnames(ccl_converter)[colnames(ccl_converter) == "Broad_ID"] <- "DepMap_ID"
ach2cvcl <- select(ccl_converter, accession_id, DepMap_ID)
lineage_converter <- unique(ccl_converter[, c("grouped_general_doid", "group_general_lineage_name")])

g2p <- read.delim("./data_munging/g2p_full_level_A_dataset_NEW_DEC_2_2018_withGroupedLineages.csv", sep = "\t", header = TRUE)
g2p[, 1:2] <- NULL
g2p <- merge(g2p, lineage_converter, by = "grouped_general_doid", all.x = TRUE)

g2p$Drug <- tolower(g2p$Drug)
g2p[g2p == -1] <- NA
g2p$Ref[g2p$Ref == "None"] <- NA
g2p$Alt[g2p$Alt == "None"] <- NA
g2p$Drug_Gene <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene), NA, paste0(g2p$Drug, "_", g2p$Gene))
g2p$Drug_Gene_Lin <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene) | is.na(g2p$group_general_lineage_name), NA, paste0(g2p$Drug, "_", g2p$Gene, "_", g2p$group_general_lineage_name))
g2p$position <- paste0("g.chr", g2p$Chromosome, ":", g2p$End, g2p$Ref, ">", g2p$Alt)
g2p$position <- ifelse(grepl("NA", g2p$position), NA, g2p$position)
g2p_filt <- subset(g2p, grepl("^CID", g2p$DrugID))

g2p_dgls <-  unique(g2p_filt$Drug_Gene_Lin[!is.na(g2p_filt$Drug_Gene_Lin)])
g2p_druggenes <- unique(g2p_filt$Drug_Gene[!is.na(g2p_filt$Drug_Gene)])
g2p_drugs <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[1]))
g2p_genes <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[2]))
g2p_lins <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[3]))
# g2p_druggenes
# g2p_drugs
# g2p_genes
# g2p_lins

# Order and genes taken from pink heatmapo from Daniel's analyses
g2p_genes <- c("EGFR", "KRAS", "ERBB2", "ERBB3", "MET", "ALK", "ROS1", "PDGFRA", "ABL1", "KIT", "PDGFRB", "PDGFB", "BRAF", "G6PD", "RET", "BRCA1", "BRCA2", "UGT1A1", "TSC1", "TSC2", "AKT1", "IDH2", "DPYD", "ESR1", "CYP19A1", "TPMT", "FLT3", "NPM1", "DNMT3A", "PML", "PTCH1", "JAK2", "MYD88", "MGMT")
```

Drug metadata and gene expression data:
```{r}
drug_meta <- read.delim("./data_munging/drug_id_drug_name_table_10_12_2018.tsv", sep = "\t", header = FALSE)
colnames(drug_meta) <- c("CID", "Drug")
drug_meta$Drug <- tolower(drug_meta$Drug)

ge_melt_18Q3 <- readRDS("./../htscreens_giant_files/ge_melt_18Q3.rds")
# ge <- filter(ge_melt_18Q3, DepMap_ID %in% ach2cvcl$DepMap_ID)
# ge <- dcast(ge[sample(nrow(ge), 5000),], CCLE_Name + DepMap_ID ~ Hugo_Symbol, value.var = "RPKM")
ge <- filter(ge_melt_18Q3, Hugo_Symbol %in% g2p_genes)
ge <- dcast(ge, CCLE_Name + DepMap_ID ~ Hugo_Symbol, value.var = "RPKM")
ge <- merge(ach2cvcl, ge, by = "DepMap_ID")
rownames(ge) <- ge$accession_id
ge[, 1:3] <- NULL
ge[] <- sapply(ge, function(x) log2(x + 0.25))
```

## CCLE

```{r}
ccle <- read.delim("./data_munging/data_drug_auc_ccle_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ccle <- merge(drug_meta, ccle, by = "CID")
ccle <- filter(ccle, Drug %in% g2p_drugs)
rownames(ccle) <- ccle$Drug
ccle$Drug <- NULL
ccle$CID <- NULL
ccle$Dataset <- NULL
ccle[] <- sapply(ccle, function(x) as.numeric(as.character(x)))
ccle <- t(ccle)

ccle <- ccle[intersect(rownames(ccle), rownames(ge)),]
ge_ccle <- ge[intersect(rownames(ccle), rownames(ge)),]

# Test only erlotinib
ccle_eqn <- data.frame(depy = I(as.matrix(ccle[,2])), indx = I(as.matrix(ge_ccle))) 
ccle_plsr1 <- plsr(depy ~ indx, data = ccle_eqn, validation = "LOO")
summary(ccle_plsr1)

par(mfrow = c(1, 2))
ccle_erlot_ncomp.onesigma <- selectNcomp(ccle_plsr1, method = "onesigma", plot = TRUE, main = "One-Sigma")
ccle_erlot_ncomp.permut <- selectNcomp(ccle_plsr1, method = "randomization", plot = TRUE, main = "Permutation")
par(mfrow = c(1, 1))
```

```{r}
plot(RMSEP(ccle_plsr1))
plot(ccle_plsr1, ncomp = ccle_erlot_ncomp.onesigma, asp = 1, line = TRUE)
plot(ccle_plsr1, plottype = "scores", comps = 1:5)
plot(ccle_plsr1, plottype = "loadings", comps = 1:5, legendpos = "topright")
plot(ccle_plsr1, plottype = "correlation", comps = 1:5)
plot(ccle_plsr1, plottype = "coef", ncomp = 1:5, legendpos = "topright")
```

## CTRP

```{r}
ctrp <- read.delim("./data_munging/data_drug_auc_ctrp_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ctrp <- merge(drug_meta, ctrp, by = "CID")
ctrp <- filter(ctrp, Drug %in% g2p_drugs)
rownames(ctrp) <- ctrp$Drug
ctrp$Drug <- NULL
ctrp$CID <- NULL
ctrp$Dataset <- NULL
ctrp[] <- sapply(ctrp, function(x) as.numeric(as.character(x)))
ctrp <- t(ctrp)

ctrp <- ctrp[intersect(rownames(ctrp), rownames(ge)),]
ge_ctrp <- ge[intersect(rownames(ctrp), rownames(ge)),]

ctrp_eqn <- data.frame(depy = I(as.matrix(ctrp[,6])), indx = I(as.matrix(ge_ctrp))) 
ctrp_plsr1 <- plsr(depy ~ indx, data = ctrp_eqn, validation = "LOO")
summary(ctrp_plsr1)
```

```{r, eval=FALSE}
ctrp_plsr <- list()
for(i in 1:ncol(ctrp)) {
  ctrp_eqn <- data.frame(depy = I(as.matrix(ctrp[,i])), indx = I(as.matrix(ge_ctrp)))
  ctrp_plsr_new <- plsr(depy ~ indx, data = ctrp_eqn, validation = "LOO")
  ctrp_plsr[[i]] <- ctrp_plsr_new
  # summary(ctrp_plsr1)
}
names(ctrp_plsr) <- colnames(ctrp)

test <- selectNcomp(ctrp_plsr[["fluorouracil"]], method = "randomization", plot = TRUE)
```
